---
title: "rnnpsycholing Binding Gender"
output:
  pdf_document: default
  html_notebook: default
---

We are looking for (1) whether the network represents stereotypical gender for purposes of reflexive pronoun agreement, (2) whether it is affected by local non-c-commanding NPs, which would indicate similarity-based interference / local coherence effects in the network's context representation.

```{r}
rm(list = ls())
library(tidyverse)
library(lme4)
library(lmerTest)
library(plotrix)

REGION_ORDER = c("Antecedent", "Relative Clause", "RC Object", "Matrix Clause", "Reflexive", "End")
REGION_EXEMPLARS = c("The surgeon", "who treated", "the sergeant / the cheerleader", "had pricked", "himself / herself", ". <eos>")
NUM_REGIONS = length(REGION_ORDER)

rename_model = function(d) {
  d %>% mutate(model=if_else(model == "gulordava", "GRNN", if_else(model == "google", "JRNN", model)))
}

d = read_csv("tests/combined_results.csv") %>%
  select(-X1) %>%
  rename_model() %>%
  group_by(model, sent_index, condition) %>%
    mutate(sent_unk=any(unk)) %>%
    ungroup() %>%
  filter(!sent_unk) %>%
  mutate(region=if_else(region == "Reflexive Mismatch" | region == "Reflexive Match", "Reflexive", region),
         region=if_else(region == "RC Object Match" | region == "RC Object Mismatch", "RC Object", region),
         region=factor(region, levels=REGION_ORDER)) %>%
  mutate(condition=if_else(condition == "refl-mismatch_rc-mismatch", "refl-mismatch_rc-match",
                   if_else(condition == "refl-mismatch_rc-match", "refl-mismatch_rc-mismatch", condition))) %>%
  separate(condition, sep="_", into=c("refl_match", "rc_match")) %>% # will throw a warning, that's ok
  separate(refl_match, sep="-", into=c("junk", "refl_match")) %>%
  separate(rc_match, sep="-", into=c("trash", "rc_match")) %>%
  select(-junk, -trash) %>%
  mutate(rc_match=if_else(is.na(rc_match), "epsilon", rc_match)) %>%
  filter(!sent_unk)

d_agg = d %>% 
  group_by(model, sent_index, region, refl_match, rc_match) %>% 
    summarise(surprisal=sum(surprisal)) %>%
    ungroup() 
  
```

### Epsilon Case

```{r}
d2 = d_agg %>%
  filter(region == "Reflexive") %>%
  filter(rc_match == "epsilon")

d2 %>%
 group_by(model, region, sent_index) %>%
    mutate(item_mean=mean(surprisal)) %>%
    ungroup() %>%
 group_by(model, region, refl_match) %>%
    summarise(m=mean(surprisal),
              s=std.error(surprisal - item_mean),
              upper=m + 1.96*s,
              lower=m - 1.96*s) %>%
    ungroup() %>%
  filter(model == "JRNN") %>%
  ggplot(aes(x=refl_match, y=m, ymin=lower, ymax=upper,fill=refl_match)) +
    ylab("Surprisal at reflexive pronoun") +
    xlab("Subject NP & reflexive gender match") +
    geom_bar(stat="identity") +
    theme(legend.position="none") +
    geom_errorbar(width=.1)
ggsave("stereotypic-gender.pdf",height=4.5,width=3.5)

```


```{r}
mj = lmer(surprisal ~ refl_match + (1|sent_index), data=filter(d2, model == "JRNN"))
summary(mj)

mg = lmer(surprisal ~ refl_match + (1|sent_index), data=filter(d2, model == "GRNN"))
summary(mg)
```

When there is no intervening clause, then a mismatched reflexive is significantly more surprising than a matched reflexive.

### Effects of Interveners 

```{r}
d2 = d_agg %>%
  filter(region == "Reflexive") %>%
  filter(rc_match != "epsilon")

d2 %>%
  filter(model == "JRNN") %>%
  group_by(sent_index) %>%
    mutate(item_mean=mean(surprisal)) %>%
    ungroup() %>%
 group_by(refl_match, rc_match) %>%
    summarise(m=mean(surprisal),
              s=std.error(surprisal - item_mean),
              upper=m + 1.96*s,
              lower=m - 1.96*s) %>%
    ungroup() %>%
  ggplot(aes(x=refl_match, y=m, ymin=lower, ymax=upper, fill=rc_match)) +
    geom_bar(stat="identity", position="dodge") +
    ylim(c(0,10)) + 
    ylab("Surprisal at reflexive pronoun") +
    xlab("Subject NP & reflexive gender match") +
    guides(fill=guide_legend(title="Intervener & reflexive gender")) +
    #theme(legend.position=c(0.4,0.9)) + 
    theme(legend.position = "top") +
    geom_errorbar(position=position_dodge(0.9), width=.1)
ggsave("binding-interveners.pdf",height=4.5,width=3.5)

```

### mismatch effects, distractor and commander

```{r}
d5 = d_agg %>%
  filter(rc_match != "epsilon", region == "Reflexive",refl_match=="mismatch")
  
m = lmer(surprisal~rc_match + (1|sent_index), data=filter(d5, model == "JRNN"))
summary(m)

d6 <- d_agg %>%
  filter(rc_match != "epsilon", region == "Reflexive",rc_match=="mismatch")

m = lmer(surprisal~refl_match + (1|sent_index), data=filter(d6, model == "JRNN"))
summary(m)

```

