---
title: "rnnpsycholing Japanese NPI (no embedded sentences)"
author: "Takashi Morita"
output:
  pdf_document:
    latex_engine: lualatex
    toc: true
    toc_depth: 3
  html_notebook: default
documentclass: ltjsarticle
---

# Introduction

We are looking for a 2x2 interaction of:

- presence vs. absence of the Japanese NPI *shika* (しか)
- affirmativeness vs. negativeness of main verb

for each of the three grammatical cases (TOP, ACC, DAT) of the *shika*-attached NP.

e.g.

- TOP
	- 渡辺-{しか,は} 家族-に 手紙-を {渡した,渡さなかった}。
	- Watanabe-{*shika*,TOP} family-DAT letter-ACC {came, didn't come}.
	- 'Only Watanabe handed letters to his family.'
	- 'Watanabe handed/didn't hand letters to his family.'
- ACC
	- 渡辺-は 家族-に 手紙-{しか,を} {渡した,渡さなかった}。
	- Watanabe-TOP family-DAT letter-{*shika*,ACC} {came, didn't.come.}
	- 'Watanabe handed only letters to his family.'
	- 'Watanabe handed/didn't hand letters to his family.'
- DAT
	- 渡辺-は 家族-{に-しか,に} 手紙-を {渡した,渡さなかった}。
	- Watanabe-TOP family-{DAT-*shika*,DAT} letter-ACC {came, didn't.come.}
	- 'Watanabe handed letters only to his family.'
	- 'Watanabe handed/didn't hand letters to his family.'

Why is this interesting?

- A grammatical sentence with *shika* must have a negative verb.
- Affirmative verbs would show significant increase in surprisal when *shika* precedes compared with its absence.



# Methods

For each pair $i$ of sentences with vs. without *shika*,
we look at their difference in surprisal of the verb (`V`) region.

\[
D_i
	 := 
		S(\texttt{V}_i \mid \texttt{shika})
		-
		S(\texttt{V}_i \mid \texttt{no-shika})
\]

\[
S(r) := - \log_2 P(r)
\]

And we perform a statistical analysis and check
if the affirmativeness vs. negativeness of the verb
have an effect on the surprisal difference $D$.


# Load data

```{r}
rm(list = ls())
library(tidyverse)
library(lme4)
library(lmerTest)
library(plotrix)


REGIONS = c('prefix', 'V', 'end')


token_based_data_path = 'jp_shika_test_sentences_unembedded_surprisal-per-token.tsv'
data_token_based = read_tsv(token_based_data_path)
# Fill the initial surprisal by 0.
data_token_based[is.na(data_token_based$surprisal),]$surprisal = 0
data_token_based$region = factor(data_token_based$region, levels=REGIONS)



data_region_based = data_token_based %>% 
	group_by(sent_index, region, shika, verb_type, shika_case) %>% 
		summarise(surprisal=sum(surprisal)) %>%
		ungroup() %>% 
	mutate(
		shika=factor(shika, levels=c("shika", "no-shika")),
		verb_type=factor(verb_type, levels=c("affirmative", "negative")),
		shika_case=factor(shika_case, levels=c("TOP", "ACC", "DAT"))
		)

# Sum coding of the variables.
contrasts(data_region_based$shika) = "contr.sum"
contrasts(data_region_based$verb_type) = "contr.sum"

# Make sure that the dataframe is sorted appropriately.
# First by case
data_region_based = data_region_based[order(data_region_based$shika_case),]
# Second by verb_type (affirmative vs. negative)
data_region_based = data_region_based[order(data_region_based$verb_type),]
# Then by sent_index
data_region_based = data_region_based[order(data_region_based$sent_index),]
```

# Visualization

```{r}
# Focus on the V (verb) region.
data_V = subset(data_region_based, region == 'V')

# Get difference in surprisal between shika vs. no-shika.
data_V_shika = subset(data_V, shika == 'shika')
data_V_no_shika = subset(data_V, shika == 'no-shika')
data_V_shika$surprisal_diff = data_V_shika$surprisal - data_V_no_shika$surprisal

# Visualize the difference in surprisal increase/dicrease between affirmative vs. negative verbs.
data_V_shika %>% 
<<<<<<< HEAD
  group_by(sent_index) %>%
    mutate(item_mean=mean(surprisal_diff)) %>%
    ungroup() %>%
	group_by(verb_type, shika_case) %>%
	  summarise(m=mean(surprisal_diff),
			s=std.error(surprisal_diff - item_mean),
=======
  
  #Error Calculation
  group_by(sent_index) %>%
   mutate(across_condition_mean = mean(surprisal_diff)) %>%
  ungroup() %>%
  mutate(item_mean = surprisal_diff - across_condition_mean) %>%
  group_by(verb_type, shika_case) %>%
    mutate(err = std.error(item_mean, na.rm=T)) %>%
  ungroup() %>%
  select(-item_mean, -across_condition_mean) %>%
  
	group_by(verb_type, shika_case) %>%
	summarise(m=mean(surprisal_diff),
	    s=mean(err),
			s2=std.error(surprisal_diff),
>>>>>>> bd50d7cfee64e8333ab0c4ebc3b0f21229b24199
			upper=m + 1.96*s,
			lower=m - 1.96*s) %>%
	  ungroup() %>%
	ggplot(aes(x=shika_case, y=m, ymin=lower, ymax=upper, fill=verb_type)) +
    ylab("Surprisal effect of shika in context") +
    xlab("Case of shika-marked NP") +
		geom_bar(stat = 'identity', position = "dodge") +
		geom_errorbar(position=position_dodge(0.9), width=.1) +
    theme(legend.position=c(0.15,0.9)) + 
    guides(fill=guide_legend(title="Verb")) 
    #annotate("text",x=1.5,y=-3,label="Unembedded")
ggsave("shika-unembedded.pdf",height=4,width=3)
```
- TOP
	- No visible increase in surprisal of the affirmative verbs.
	- Visible decrease in surprisal of the negative verbs.
- ACC
	- Greatest increase in surprisal of the affirmative verbs.
	- Small decrease in surprisal of the negative verbs.
- DAT
	- Visible increase in surprisal of the affirmative verbs.
	- Greatest decrease in surprisal of the negative verbs.

# Regressions

## TOP

```{r}
sub_data = subset(data_V_shika, shika_case == 'TOP')

m = lmer(
		surprisal_diff
			~ verb_type
				+ (1 | sent_index)
		,
		data=sub_data
		)
summary(m)
```

- Statistically significant effect of the affirmativeness vs. negativeness of verbs.


## ACC

```{r}
sub_data = subset(data_V_shika, shika_case == 'ACC')

m = lmer(
		surprisal_diff
			~ verb_type
				+ (1 | sent_index)
		,
		data=sub_data
		)
summary(m)
```


- Statistically significant effect of the affirmativeness vs. negativeness of verbs.
- Greater effect than TOP.

## DAT

```{r}
sub_data = subset(data_V_shika, shika_case == 'DAT')

m = lmer(
		surprisal_diff
			~ verb_type
				+ (1 | sent_index)
		,
		data=sub_data
		)
summary(m)
```

- Statistically significant effect of the affirmativeness vs. negativeness of verbs.
- Greater effect than TOP, similar to ACC.

