---
title: "rnnpsycholing Japanese NPI (sentences with embedding and matrix shika)"
author: "Takashi Morita"
output:
  pdf_document:
    latex_engine: lualatex
    toc: true
    toc_depth: 3
  html_notebook: default
documentclass: ltjsarticle
---

# Introduction

We are looking for a 2x2x2 interaction of:

- presence vs. absence of the Japanese NPI *shika* (しか) in the main clause.
- affirmativeness vs. negativeness of main verb
- affirmativeness vs. negativeness of embedded verb

for each of the three grammatical cases (TOP, DAT) of the *shika*-attached NP.

e.g.

- TOP
	- 佐藤-{しか, は} 社長-が パーティ-に 友人-を {呼んだ, 呼ばなかった} と {思った,思わなかった}。
	- Sato-{*shika*, TOP} CEO-NOM party-DAT friend-ACC {invited, didn't invite} that {thought, didn't think}.
- DAT
	- 同僚-{にしか, に} 佐藤は 社長-が パーティ-に 友人-を {呼んだ, 呼ばなかった} と {思った,思わなかった}。
	- colleague-{DAT-*shika*, DAT} Sato-TOP CEO-NOM party-DAT friend-ACC {invited, didn't invite} that {thought, didn't think}.

Why is this interesting?

1. A grammatical sentence with *shika* in the main clause must have a negative main verb.
	- A significant increase in surprisal of the affirmative main verbs must be predicted by the LSTM conditioned on the presence of *shika* if the learning is successful.
2. Negation of the embedded verb does not satisfy the *shika*'s grammatical condition.
	- No significant increase in surprisal of the affirmative embedded verbs given *shika* is expected for a successful learner.
	- Nor significant interaction between the main and embedded verbs given *shika* is expected for a successful learner.


# Load data


```{r}
rm(list = ls())
library(tidyverse)
library(lme4)
library(lmerTest)
library(plotrix)

REGIONS = c('main_prefix', 'embedded_prefix', 'embedded_V', 'complementizer', 'main_V', 'end')

token_based_data_path = 'jp_shika_test_sentences_embedded_shika-in-main_surprisal-per-token.tsv'
data_token_based = read_tsv(token_based_data_path)
# Fill the initial surprisal by 0.
data_token_based[is.na(data_token_based$surprisal),]$surprisal = 0
data_token_based$region = factor(data_token_based$region, levels=REGIONS)

data_region_based = data_token_based %>% 
	group_by(sent_index, region, shika, embed_V, main_V, shika_case) %>% 
		summarise(surprisal=sum(surprisal)) %>%
		ungroup() %>% 
	mutate(
		shika=factor(shika, levels=c("shika", "no-shika")),
		embed_V=factor(embed_V, levels=c("affirmative", "negative")),
		main_V=factor(main_V, levels=c("affirmative", "negative")),
		shika_case=factor(shika_case, levels=c("TOP", "DAT"))
		)

# Sum coding of the variables.
contrasts(data_region_based$shika) = "contr.sum"
contrasts(data_region_based$embed_V) = "contr.sum"
contrasts(data_region_based$main_V) = "contr.sum"

# Make sure that the dataframe is sorted appropriately.
# First by embed_V (affirmative vs. negative)
data_region_based = data_region_based[order(data_region_based$embed_V),]
# Then by main_V
data_region_based = data_region_based[order(data_region_based$main_V),]
# finally by sent_index
data_region_based = data_region_based[order(data_region_based$sent_index),]
```

# Embedded verb region

## Visualization

```{r}
# Focus on the V (verb) region.
data_V = subset(data_region_based, region == 'embedded_V')

# Get difference in surprisal between shika vs. no-shika.
data_V_shika = subset(data_V, shika == 'shika')
data_V_no_shika = subset(data_V, shika == 'no-shika')
data_V_shika$surprisal_diff = data_V_shika$surprisal - data_V_no_shika$surprisal

# Visualize the difference in surprisal increase/dicrease between affirmative vs. negative verbs.
data_V_shika %>% 
<<<<<<< HEAD
  group_by(sent_index) %>%
    mutate(item_mean=mean(surprisal_diff)) %>%
    ungroup() %>%
	group_by(embed_V, shika_case) %>%
	summarise(m=mean(surprisal_diff),
			s=std.error(surprisal_diff - item_mean),
=======
  
  #Error Calculation
  group_by(sent_index) %>%
   mutate(across_condition_mean = mean(surprisal_diff)) %>%
  ungroup() %>%
  mutate(item_mean = surprisal_diff - across_condition_mean) %>%
  group_by(embed_V, shika_case) %>%
    mutate(err = std.error(item_mean, na.rm=T)) %>%
  ungroup() %>%
  select(-item_mean, -across_condition_mean) %>%
  
	group_by(embed_V, shika_case) %>%
	summarise(m=mean(surprisal_diff),
			s2=std.error(surprisal_diff),
			s=mean(err),
>>>>>>> bd50d7cfee64e8333ab0c4ebc3b0f21229b24199
			upper=m + 1.96*s,
			lower=m - 1.96*s) %>%
	ungroup() %>%
	ggplot(aes(x=shika_case, y=m, ymin=lower, ymax=upper, fill=embed_V)) +
    ylab("Surprisal effect of shika in context") +
    xlab("Case of shika-marked NP") +
		geom_bar(stat = 'identity', position = "dodge") +
		geom_errorbar(position=position_dodge(0.9), width=.1) +
    #theme(legend.position=c(0.1,0.4)) +
    theme(legend.position="none") +
    #annotate("text",x=1.5,y=-0.8,label="Complex, shika in matrix") +
    guides(fill=guide_legend(title="Verb")) 
ggsave("complex-shika-in-matrix.pdf",height=4,width=3)
```

## Regressions
### TOP

```{r}
sub_data = subset(data_V_shika, shika_case == 'TOP')

m = lmer(
		surprisal_diff
			~ embed_V
				+ (1 | sent_index)
		,
		data=sub_data
		)
summary(m)
```

- Significant negative effect of embed_V (affirmativeness = 1).
	- Negative verbs cause more 


### DAT

```{r}
sub_data = subset(data_V_shika, shika_case == 'DAT')

m = lmer(
		surprisal_diff
			~ embed_V
				+ (1 | sent_index)
		,
		data=sub_data
		)
summary(m)
```

- Significant positive effect of embed_V (affirmativeness = 1).

# Main verb region

## Visualization

```{r}
# Focus on the V (verb) region.
data_V = subset(data_region_based, region == 'main_V')

# Get difference in surprisal between shika vs. no-shika.
data_V_shika = subset(data_V, shika == 'shika')
data_V_no_shika = subset(data_V, shika == 'no-shika')
data_V_shika$surprisal_diff = data_V_shika$surprisal - data_V_no_shika$surprisal


# Visualize the difference in surprisal increase/dicrease between affirmative vs. negative verbs.
data_V_shika %>% 
<<<<<<< HEAD
  group_by(sent_index) %>%
    mutate(item_mean=mean(surprisal_diff)) %>%
    ungroup() %>%
	group_by(embed_V, main_V, shika_case) %>%
	summarise(m=mean(surprisal_diff),
			s=std.error(surprisal_diff-item_mean),
=======
  
  #Error Calculation
  group_by(sent_index) %>%
   mutate(across_condition_mean = mean(surprisal_diff)) %>%
  ungroup() %>%
  mutate(item_mean = surprisal_diff - across_condition_mean) %>%
  group_by(embed_V, shika_case, main_V) %>%
    mutate(err = std.error(item_mean, na.rm=T)) %>%
  ungroup() %>%
  select(-item_mean, -across_condition_mean) %>%
  
	group_by(embed_V, main_V, shika_case) %>%
	summarise(m=mean(surprisal_diff),
			s2=std.error(surprisal_diff),
			s=mean(err),
>>>>>>> bd50d7cfee64e8333ab0c4ebc3b0f21229b24199
			upper=m + 1.96*s,
			lower=m - 1.96*s) %>%
	ungroup() %>%
	ggplot(aes(x=shika_case, y=m, ymin=lower, ymax=upper, fill=embed_V:main_V)) +
    ylab("Surprisal effect of shika in context") +
    xlab("Case of shika-marked NP") +
    theme(legend.position=c(0.3,0.2)) +
		geom_bar(stat = 'identity', position = "dodge") +
		geom_errorbar(position=position_dodge(0.9), width=.1)
ggsave("complex-shika-in-matrix-matrixverb.pdf",height=4,width=4)
```

- TOP
	- Increas in surprisal in every condition.
- DAT
	- Small but expected signs of changes.

## Regressions

### TOP

```{r}
sub_data = subset(data_V_shika, shika_case == 'TOP')

m = lmer(
		surprisal_diff
			~ embed_V * main_V
				+ (embed_V + main_V | sent_index)
		,
		data=sub_data
		)
summary(m)
```

- No significant effect of embed_V (affirmativeness = 1).
- Significant negative effect of main_V (affirmativeness = 1).
- Significant negative interaction.

### DAT

```{r}
sub_data = subset(data_V_shika, shika_case == 'DAT')

m = lmer(
		surprisal_diff
			~ embed_V * main_V
				+ (embed_V + main_V | sent_index)
		,
		data=sub_data
		)
summary(m)
```

- Significant negative effect of embed_V (affirmativeness = 1).
- Significant positive effect of main_V (affirmativeness = 1).
- Significant positive interaction.