---
title: "Analysis 2 English NPI Data"
output: html_notebook
---

---
title: "rnnpsycholing Binding Gender"
output:
  pdf_document: default
  html_notebook: default
---

We are looking for (1) whether the network represents stereotypical gender for purposes of reflexive pronoun agreement, (2) whether it is affected by local non-c-commanding NPs, which would indicate similarity-based interference / local coherence effects in the network's context representation.

```{r}
rm(list = ls())
library(tidyverse)
library(lme4)
library(lmerTest)
library(plotrix)

REGION_ORDER = c("Antecedent", "Relative Clause", "Aux", "Ever", "Verb", "Any", "Noun", "Continuation", "Conclusion")
REGION_EXEMPLARS = c("No bill", "that the/no senator likes", "has", "ever", "gotten", "any", "support in the senate", ". <eos>")
NUM_REGIONS = length(REGION_ORDER)

rename_model = function(d) {
  d %>% mutate(model=if_else(model == "gulordava", "GRNN", if_else(model == "google", "JRNN", model)))
}

d = read_csv("tests/combined_results.csv") %>%
  select(-X1) %>%
  rename_model() %>%
  group_by(model, sent_index, condition) %>%
    mutate(sent_unk=any(unk)) %>%
    ungroup() %>%
  filter(!sent_unk) %>%
  mutate(region=if_else(region == "Non-Negative Licensor" | region == "Negative Licensor", "Licensor", region),
         region=if_else(region == "Distractor RC" | region == "RC", "Relative Clause", region),
         region=factor(region, levels=REGION_ORDER)) %>%
  separate(condition, sep="_", into=c("commander", "distractor", "ever", "any")) %>% # will throw a warning, that's ok
  filter(!sent_unk) %>%
  mutate(distractor = factor(ifelse(distractor == "nprc", "neg", "pos")))


d_agg = d %>% 
  group_by(model, sent_index, region, commander, distractor, ever, any) %>% 
    summarise(surprisal=sum(surprisal)) %>%
    ungroup() 
  
```

### Effects of Any

```{r}
d2 = d_agg %>%
  filter(region == "Any") %>%
  filter(ever == "never", any == "any")

d2 %>%
  group_by(model, sent_index) %>%
    mutate(item_mean=mean(surprisal)) %>%
    ungroup() %>%
 group_by(model, commander, distractor) %>%
    summarise(m=mean(surprisal),
              s=std.error(surprisal - item_mean),
              upper=m + 1.96*s,
              lower=m - 1.96*s) %>%
    ungroup() %>%
  ggplot(aes(x=commander, y=m, ymin=lower, ymax=upper, fill=distractor)) +
    geom_bar(stat="identity", position="dodge") +
    ylab("Surprisal at NPI") +
    xlab("Matrix Licensor Polarity") +
    guides(fill=guide_legend(title="Intervener Polarity")) +
    theme(legend.position="top") + 
    geom_errorbar(position=position_dodge(0.9), width=.1) +
    facet_wrap(~model) +
    ggtitle("Surprisal of 'Any'")
ggsave("npi-interveners-any.pdf",height=4.5,width=3.5)

```

### mismatch effects, distractor and commander

```{r}
d5 = d_agg %>%
  filter(region == "Any", ever=="never", any == "any")
  
m = lmer(surprisal~commander + (1|sent_index), data=filter(d5, model == "JRNN"))
summary(m)
m = lmer(surprisal~commander + (1|sent_index), data=filter(d5, model == "GRNN"))
summary(m)

d6 <- d_agg %>%
  filter(region == "Any", ever=="never", any == "any")

m = lmer(surprisal~distractor + (1|sent_index), data=filter(d6, model == "JRNN"))
summary(m)
m = lmer(surprisal~distractor + (1|sent_index), data=filter(d6, model == "GRNN"))
summary(m)
```

```{r}
d2 = d_agg %>%
  filter(region == "Ever") %>%
  filter(ever == "ever", any == "nany")

d2 %>%
  group_by(model, sent_index) %>%
    mutate(item_mean=mean(surprisal)) %>%
    ungroup() %>%
 group_by(model, commander, distractor) %>%
    summarise(m=mean(surprisal),
              s=std.error(surprisal - item_mean),
              upper=m + 1.96*s,
              lower=m - 1.96*s) %>%
    ungroup() %>%
  ggplot(aes(x=commander, y=m, ymin=lower, ymax=upper, fill=distractor)) +
    geom_bar(stat="identity", position="dodge") +
    ylab("Surprisal at NPI") +
    xlab("Matrix Licensor Polarity") +
    guides(fill=guide_legend(title="Intervener Polarity")) +
    #theme(legend.position=c(0.22,0.9)) + 
    theme(legend.position="top") + 
    geom_errorbar(position=position_dodge(0.9), width=.1) +
    facet_wrap(~model) +
    ggtitle("Surprisal of 'Ever'")
ggsave("npi-interveners-ever.pdf",height=4.5,width=3.5)

```

```{r}
d7 = d_agg %>%
  filter(region == "Ever", ever=="ever", any == "nany")
  
m = lmer(surprisal~commander + (1|sent_index), data=filter(d5, model == "JRNN"))
summary(m)
m = lmer(surprisal~commander + (1|sent_index), data=filter(d5, model == "GRNN"))
summary(m)

d8 <- d_agg %>%
  filter(region == "Ever", ever=="ever", any=="nany")

m = lmer(surprisal~distractor + (1|sent_index), data=filter(d6, model == "JRNN"))
summary(m)
m = lmer(surprisal~distractor + (1|sent_index), data=filter(d6, model == "GRNN"))
summary(m)
```